---
title: Example Time To Event (Survival) Analysis using visR
author: ZERO
date: '2020-05-04'
slug: example-time-to-event-survival-analysis-using-visr
categories:
  - Tools
tags:
  - clinical
subtitle: ''
summary: ''
authors: []
lastmod: '2020-05-04T18:18:49+08:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---


## setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = TRUE,      # Output code chunks
    message = FALSE,  # Toggle off message output 
    warning = FALSE) 

knitr::opts_knit$set(root.dir = usethis::proj_path())
library(docknitr)

# libraries used in report
library(knitr)
library(kableExtra)
library(tidyverse)

theme_set(theme_light(base_family = "Avenir"))
```

## load pkgs
```{r}
library(visR)
library(survival)
library(gt)
```


## set var
```{r}
# Constants
DATASET <- paste0("NCCTG Lung Cancer Dataset (from survival package ", 
                  packageVersion("survival"), ")")

#attrition_chart_fn <- "data/attrition_diagram.svg"
# Globql formatting options
options(digits = 3)

# Global table settings 
options(DT.options = list(pageLength = 10, 
                          language = list(search = 'Filter:'), 
                          scrollX = TRUE))
```

## load data

```{r}
data(lung)

lung_cohort <- lung %>% 
  rename(ECOG = ph.ecog,
         Karnofsky = ph.karno,
         institution = inst
         ) %>% 
  mutate(patid = paste0("Pat", row_number()),
         institution = factor(institution),
         sex = factor(if_else(sex == 1, "male", "female")),
         ECOG = factor(case_when(ECOG == 0 ~ "0 asymptomatic",
                          ECOG == 1 ~ "1 ambulatory",
                          ECOG == 2 ~ "2 in bed less than 50% of day",
                          ECOG == 3 ~ "3 in bed more than 50% of day",
                          ECOG == 4 ~ "4 bedbound",
                          ECOG == 5 ~ "5 dead")),
         dx_age_group = factor(case_when(age < 30 ~ "< 30y",
                                         age >= 30 & age <= 50 ~ "30-50y",
                                         age > 50 & age <= 70 ~ "51-70y",
                                         age > 70 ~ "> 70y"),
                               levels=c("< 30y", "30-50y", "51-70y", "> 70y")),
    status = if_else(status == 1, 0, 1)) %>% 
    select(-meal.cal, -pat.karno)
```

## EDA
```{r}
skimr::skim(lung_cohort)
```

## follow chart Cohort Selection

For this exercise we will only include patients with (1) ECOG available (2) and non-missing weight-loss data in our analysis

### Attrition Table

过滤表

```{r}
cohort_attrition <- vr_attrition_table(
  data = lung_cohort,
  criteria_descriptions = c("1. ECOG available", 
                            "2. Weight loss data available", 
                            "3. Non-missing vital status",
                            "4. Positive follow up time"),
  criteria_conditions   = c("!is.na(ECOG)",
                            "!is.na(wt.loss)",
                            "!is.na(status)",
                            "time >= 0"),
  subject_column_name   = 'patid')

 
vr_render_table(data = cohort_attrition, 
                title = "Attrition Table", 
                caption = "Summary of samples fulfilling inclusion/exclusion criteria", 
                datasource = DATASET, 
                engine = "gt")
```


### flow-chart diagram

```{r}
complement_descriptions <- c(                                                   
    "Having exclusion criterion: non ECOG",                                            
    "Having exclusion criterion: missing weight loss ",
    "Having exclusion criterion: missing vital status",
    "Having exclusion criterion: negative follow up time"
)

# Create attrition flowchart
#attrition_chart_fn <- "attrition_diagram.svg"
# attrition_flow <- vr_attrition(
#   N_array = cohort_attrition$`Remaining N`,
#   descriptions = cohort_attrition$Criteria,
#   complement_descriptions = complement_descriptions,
#   output_path = attrition_chart_fn)

# knitr::include_graphics(attrition_chart_fn)
```

### analysis data

```{r}
lung_cohort <- lung_cohort %>% 
  filter(!is.na(ECOG),
         !is.na(wt.loss),
         !is.na(status),
         time >= 0)
```

## Table 1

Cohort Overview

if the there are systematic differences between the chemotherapy and hormone therapy arms.



```{r}
# some modifications to the table for nice printing of categories
lung_cohort_tab1 <- lung_cohort %>% 
  mutate(status = factor(case_when(status == 0 ~ "Alive/Censored",
                                   status == 1 ~ "Dead",
                                   is.na(status) ~ "Missing"))) %>% 
  filter(!is.na(time) & !is.na(age) & !is.na(Karnofsky)) %>% 
  select(-institution)

# visR table1 convenience function
vr_table_one(
  data = lung_cohort_tab1,
  title = "Overview over Lung Cancer patients",
  caption = "Baseline characteristics of study cohort stratified by treatment type",
  datasource = DATASET,
  groupCols = c("sex")
)
```



## Kaplan-Meier Curve

```{r}
surv_eq <- "survival::Surv(time, status) ~ sex"

vr_kaplan_meier(
    lung_cohort,
    equation = surv_eq,
    data_source = DATASET, 
    time_unit = "days",
    title = "Comparison of survival in male and female lung cancer patients"
)
```

### Summary of Kaplan Meier Curve

```{r}
km_summary <- vr_kaplan_meier_summary(lung_cohort, equation = surv_eq)

km_summary[[1]] %>%
  vr_render_table("Overall Survival",
                  "Median survival times in days for each strata",
                  DATASET)


km_summary[[2]] %>%
  vr_render_table("Equality between Strata",
                  "Summary table with test of equality over strata",
                  DATASET)
```






